Matching Bullets

### 2015-12-18 imaging-paper.Rnw
ratios included on densities

\begin{figure}[hbtp]
<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='\\linewidth'>>=
library(RColorBrewer)
library(ggplot2)
library(scales)
library(dplyr)
library(x3pr)
library(x3prplus)
library(grid)
library(gridExtra)
library(zoo)
library(tidyr)
library(rpart)
library(rpart.plot)
library(xtable)

@@

@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Densities of maximum CMS for Known Matches (KM) and Known Non Matches (KNM).}
<<cms-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + 
    geom_density(aes(x=cms, fill=match), alpha=0.5) +
  xlab("maximum CMS") +
  scale_fill_brewer("", palette="Paired", labels=c("KNM", "KM")) +
  theme_bw() + 
  theme(legend.position=c(1,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) 
@
\end{subfigure}
\caption{\label{fig:cms}Overview of the maximal number of consecutive maximal striae compared to matches.}
\end{figure}

@@

For smaller numbers of CMS, the percentage of false positives quickly increases. However, if we take other features of the image into account, we can increase the number of correct matches considerably: a high amount of cross-correlation between the two cross sections (see Figure~\ref{fig:ccfs-match}) is indicative for a match; e.g.\  in the Hamby study only known matches have a cross-correlation of 0.75 or higher. There are 97 land-to-land comparisons with a cross-correlation of above 0.75.   
\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:distr.dist}Relationship between maximal CMS and distances between cross sections.}
<<distr.dist, echo=FALSE, fig.width=8, fig.height=4, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + 
    geom_jitter(aes(x=factor(cms), y=distr.dist, colour=match), alpha=0.5) +
   facet_wrap(~match, labeller="label_both") + xlab("maximum CMS") + 
  ylab("Distance between cross sections") +
  scale_colour_brewer(palette="Paired") +
  theme_bw() + 
  theme(legend.position="none") +
    xlab("maximum CMS")

@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
<<crosscut-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + 
    geom_density(aes(x=distr.dist, fill=match), alpha=0.5) +
  xlab("Distances between cross sections") +
  scale_fill_brewer("", palette="Paired", labels=c("KNM", "KM")) +
  theme_bw() + 
  theme(legend.position=c(1,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) 
@
\end{subfigure}
\caption{Relationship between distances between cross sections. A small distance between cross sections is indicative of matches between lands.}
\end{figure}

\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:ccfs-match}Relationship between maximal CMS and cross-correlation between cross sections.}
<<ccfs-match, echo=FALSE, fig.width=8, fig.height=4, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + theme_bw() + 
  geom_jitter(aes(x=factor(cms), y=ccf, colour=match), alpha=.5) + 
  facet_wrap(~match, labeller="label_both") + 
    scale_colour_brewer(palette="Paired") +
  xlab("maximum CMS") + ylab("Cross-correlation") +
  theme(legend.position="none")
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
<<ccf-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + 
    geom_density(aes(x=ccf, fill=match), alpha=0.5) +
  xlab("Cross-correlation between cross sections") +
  scale_fill_brewer("", palette="Paired", labels=c("KNM", "KM")) +
  theme_bw() + 
  theme(legend.position=c(1,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) 
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Ratio of probabilities of $P(ccf > x \mid KM) / P(ccf > x \mid KNM)$.}
<<ccf-ratio, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
x <- seq(0,1, by=0.001)
library(sm)
km <- sm.density(bstats$ccf[bstats$match], display="none", eval.points=x)
knm <- sm.density(bstats$ccf[!bstats$match], display="none", eval.points=x)
dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)

@@

ggplot() +
  geom_polygon(aes(x=x, y=estimate, fill=match, colour=match), 
               data=densities, alpha=0.5) +
  xlab("Cross-correlation between cross sections") +
  scale_fill_brewer("", palette="Paired", labels=c("KNM", "KM")) +
  scale_colour_brewer("", palette="Paired", labels=c("KNM", "KM")) +
  theme_bw() + 
  ylab("Density estimate") +
  theme(legend.position=c(.975,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) +
  geom_line(aes(x, pmin(5, kmtail/knmtail)), data=dframe) 
@
\end{subfigure}
\caption{A large cross-correlation between cross sections is indicative of matches between lands.}
\end{figure}

>>>>

\begin{figure}[hbtp]
<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='\\linewidth'>>=
library(RColorBrewer)
library(ggplot2)
library(scales)
library(dplyr)
library(x3pr)
library(x3prplus)
library(grid)
library(gridExtra)
library(zoo)
library(tidyr)
library(rpart)
library(rpart.plot)
library(xtable)
library(sm)

@@

@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Densities of maximum CMS for Known Matches (KM) and Known Non Matches (KNM).}
<<cms-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
x <- seq(0,29, by=0.03)
km <- sm.density(bstats$cms[bstats$match], display="none", eval.points=x)
knm <- sm.density(bstats$cms[!bstats$match], display="none", eval.points=x)
dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)

densities <- data.frame(
  match = rep(c("KM", "KNM"), each=(length(x)+2)),
  x = rep(c(0, x, 29), times = 2),
  estimate = c(0, km$estimate, 0, 0, knm$estimate,0))

ggplot() +
  geom_polygon(aes(x=x, y=estimate, fill=match, group=match), colour="black",
               data=densities, alpha=0.5) +
xlab("maximum CMS") +
  scale_fill_brewer("", palette="Paired") +
  ylab("Density estimates") +
  theme_bw() + 
  theme(legend.position=c(1,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) +
  geom_line(aes(x=x, y=pmin(0.2, kmtail/knmtail/10)), data=dframe)

# ggplot(data=bstats) + 
#     geom_density(aes(x=cms, fill=match), alpha=0.5) +
#   xlab("maximum CMS") +
#   scale_fill_brewer("", palette="Paired") +
#   theme_bw() + 
#   theme(legend.position=c(1,1), legend.justification=c(1,1),
#         legend.background = element_rect(fill=alpha('white', 0.4)))

@
\end{subfigure}
\caption{\label{fig:cms}Overview of the maximal number of consecutive maximal striae compared to matches.}
\end{figure}

@@

For smaller numbers of CMS, the percentage of false positives quickly increases. However, if we take other features of the image into account, we can increase the number of correct matches considerably: a high amount of cross-correlation between the two cross sections (see Figure~\ref{fig:ccfs-match}) is indicative for a match; e.g.\  in the Hamby study only known matches have a cross-correlation of 0.75 or higher. There are 97 land-to-land comparisons with a cross-correlation of above 0.75.   
\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:distr.dist}Relationship between maximal CMS and distances between cross sections.}
<<distr.dist, echo=FALSE, fig.width=8, fig.height=4, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + 
    geom_jitter(aes(x=factor(cms), y=distr.dist, colour=match), alpha=0.5) +
   facet_wrap(~match, labeller="label_both") + xlab("maximum CMS") + 
  ylab("Distance between cross sections") +
  scale_colour_brewer(palette="Paired") +
  theme_bw() + 
  theme(legend.position="none") +
    xlab("maximum CMS")
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
<<crosscut-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + 
    geom_density(aes(x=distr.dist, fill=match), alpha=0.5) +
  xlab("Distances between cross sections") +
  scale_fill_brewer("", palette="Paired") +
  theme_bw() + 
  theme(legend.position=c(1,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4)))

x <- seq(0,20, by=0.02)
km <- sm.density(bstats$distr.dist[bstats$match], display="none", eval.points=x)
knm <- sm.density(bstats$distr.dist[!bstats$match], display="none", eval.points=x)
dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)

densities <- data.frame(
  match = rep(c("KM", "KNM"), each=(length(x)+2)),
  x = rep(c(0, x, 29), times = 2),
  estimate = c(0, km$estimate, 0, 0, knm$estimate,0))

ggplot() +
  geom_polygon(aes(x=x, y=estimate, fill=match, group=match), colour="black",
               data=densities, alpha=0.5) +
  xlab("Distances between cross sections") +
  scale_fill_brewer("", palette="Paired") +
  ylab("Density estimates") +
  theme_bw() + 
  theme(legend.position=c(1,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) +
  geom_line(aes(x=x, y=pmin(0.2, kmtail/knmtail/10)), data=dframe)

@
\end{subfigure}
\caption{Relationship between distances between cross sections. A small distance between cross sections is indicative of matches between lands.}
\end{figure}

\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:ccfs-match}Relationship between maximal CMS and cross-correlation between cross sections.}
<<ccfs-match, echo=FALSE, fig.width=8, fig.height=4, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + theme_bw() + 
  geom_jitter(aes(x=factor(cms), y=ccf, colour=match), alpha=.5) + 
  facet_wrap(~match, labeller="label_both") + 
    scale_colour_brewer(palette="Paired") +
  xlab("maximum CMS") + ylab("Cross-correlation") +
  theme(legend.position="none")
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
<<ccf-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + 
    geom_density(aes(x=ccf, fill=match), alpha=0.5) +
  xlab("Cross-correlation between cross sections") +
  scale_fill_brewer("", palette="Paired") +
  theme_bw() + 
  theme(legend.position=c(1,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) 
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Ratio of probabilities of $P(ccf > x \mid KM) / P(ccf > x \mid KNM)$.}
<<ccf-ratio, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
x <- seq(0,1, by=0.001)
library(sm)
km <- sm.density(bstats$ccf[bstats$match], display="none", eval.points=x)
knm <- sm.density(bstats$ccf[!bstats$match], display="none", eval.points=x)
dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)

@@

ggplot() +
  geom_polygon(aes(x=x, y=estimate, fill=match), colour="black",
               data=densities, alpha=0.5) +
  xlab("Cross-correlation between cross sections") +
  scale_fill_brewer("", palette="Paired") +
#  scale_colour_brewer("", palette="Paired", labels=c("KNM", "KM")) +
  theme_bw() + 
  ylab("Density estimate") +
  theme(legend.position=c(.975,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) +
  geom_line(aes(x, pmin(5, kmtail/knmtail)), data=dframe) 
@
\end{subfigure}
\caption{A large cross-correlation between cross sections is indicative of matches between lands.}
\end{figure}
